---
title: "SIG07 Interaction Viz"
author: "Eric Y. Wang"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    html_preview: false
    mathjax: default
  html_notebook:
    toc: true
    toc_float: true
    mathjax: default
---

```{r setup}
library(tidyverse)
library(ggrepel)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("functions/plotting_fxns.R")
theme_set(theme_Publication())
```

## Goals

Visualize and examine the interaction effects as determined by gene-level Gamma
Poisson GLM modeling. As a reminder, the GLM is expressed as follows:

$$
\log(\mu_i) = \beta_0 + \beta_1 (\text{ligand1}_i) + \beta_2 (\text{ligand2}_i) + \beta_3 (\text{ligand1}_i \cdot \text{ligand2}_i) + \beta_4 (\text{group}_i) + \beta_5 (\text{percent.mito}_i) + \beta_6 (\text{s.score}_i) + \beta_7 (\text{g2m.score}_i) + \log(\text{total_counts}_i)
$$

Where:

-   $\mu_i$: The expected counts for observation $i$.
-   $\beta_0$: The intercept term (baseline level of expression).
-   $\beta_1$ to $\beta_7$: Coefficients representing the effect sizes of the
    predictors.
-   $\text{ligand1}_i$: The value of `ligand1` for observation $i$.
-   $\text{ligand2}_i$: The value of `ligand2` for observation $i$.
-   $\text{ligand1}_i \cdot \text{ligand2}_i$: The interaction effect between
    `ligand1` and `ligand2`.
-   $\text{group}_i$: The biological group variable (different mice).
-   $\text{percent.mito}_i$: The percentage of mitochondrial RNA for observation
    $i$.
-   $\text{s.score}_i$: The S-phase cell cycle score for observation $i$.
-   $\text{g2m.score}_i$: The G2/M-phase cell cycle score for observation $i$.
-   $\text{total_counts}_i$: The scaling factor for sequencing depth.

## Import Data

Here, the LFCs are calculated from the GLM by performing QLRTs for the
interaction term and each individual ligand. The DEG file is just for reference
and it is from a simple wilcoxon test performed in scanpy.

```{r}
# import LFC from glm model
interactionsLfc <- read_csv("analysis_outs/glmGamPoi_interaction_deg.csv")
interSig <- interactionsLfc %>% filter(adj_pval < 0.1)
singlesLfc <- read_csv("analysis_outs/glmGamPoi_singles_deg.csv")
# import DEGs from scanpy
degWilcox <- read_csv("analysis_outs/deg_wilcoxon.csv")[-1]
```

```{r}
interSig$interaction %>% unique()
```

## Helper functions

```{r make_xy_plot}
make_xy_plot <- function(ligand_pair, interactions, singles,
                         p_val_cutoff=0.1, lfc_cutoff=0.25){
  # Function to create an XY plot comparing log-fold changes (LFC) for two ligands in a pair.
  # Inputs:
  # - ligand_pair: A string specifying the ligand pair, e.g., "IL4_IFNA".
  # - interactions: Dataframe containing interaction data with columns `name`, `lfc`, `adj_pval`, `interaction`.
  # - singles: Dataframe containing single ligand data with columns `name`, `lfc`, `adj_pval`, `single_ligand`.
  # - p_val_cutoff: P-value cutoff for filtering significant interactions (default = 0.1).
  # - lfc_cutoff: Log-fold change cutoff for filtering significant interactions (default = 0.25).

  # Split the ligand pair (e.g., "IL4_IFNA") into individual ligands
  ligands <- str_split_1(ligand_pair, pattern = "_")
  
  # Filter and process the singles data
  singles <- singles %>%
    filter(single_ligand %in% ligands) %>%
    filter(interaction == ligand_pair) %>%
    select(name, adj_pval, lfc, single_ligand) %>%
    pivot_wider(
      names_from = single_ligand,         
      values_from = c(adj_pval, lfc),       
      names_sep = "_"                       
    )

  # Filter and process the interaction data
  interactions <- interactions %>%
    filter(adj_pval < p_val_cutoff) %>%
    filter(interaction == ligand_pair) %>%
    filter(abs(lfc) > lfc_cutoff) %>%
    left_join(singles, by = "name")

  # Define vlag palette
  vlag_palette <- c("#B2182B", "#D6604D", "#F4A582", "#FDDBC7", "#FFFFFF",
                    "#D1E5F0", "#92C5DE", "#4393C3", "#2166AC") %>% rev()

  # Set the limits for the log-fold change scale, rounded up to one decimal place
  lfc_max <- ceiling(max(abs(interactions$lfc)) * 10) / 10

  # plot
  ggplot(interactions, aes(x=.data[[paste0("lfc_",ligands[1])]],
                           y=.data[[paste0("lfc_",ligands[2])]],
                           fill=lfc,
                           size = -log10(adj_pval))) +
    geom_point(shape=21, alpha=0.7) +
    scale_fill_gradientn(
      colors = vlag_palette,
      limits = c(-lfc_max, lfc_max)
    ) +
    geom_hline(yintercept=0, linetype="dashed") +
    geom_vline(xintercept=0, linetype="dashed") +
    theme(aspect.ratio=1)
}
```

```{r plot_lfc_pair}
plot_lfc_pair <- function(genes, ligand_pair, deg_data) {
  # Function to plot log-fold changes (LFC) for interaction and single ligands given a vector of genes
  
  # Split the ligand pair (e.g., "IL4_IFNA") into individual ligands (e.g., "IL4" and "IFNA")
  ligands <- str_split_1(ligand_pair, pattern = "_")
  
  # Filter and process the input differential expression (DEG) data
  deg_data %>%
    filter(
      # Select only rows where the group matches the ligands and their interaction
      group %in% c(
        paste0(ligands[1], "_linker"),  # Single ligand 1 with "_linker"
        paste0(ligands[2], "_linker"),  # Single ligand 2 with "_linker"
        ligand_pair                     # Interaction of both ligands
      )
    ) %>%
    # Convert the `group` column into a factor to ensure consistent ordering
    mutate(group = factor(
      group,
      levels = c(
        paste0(ligands[1], "_linker"),  # First ligand "_linker"
        paste0(ligands[2], "_linker"),  # Second ligand "_linker"
        ligand_pair                     # Combined ligand interaction
      )
    )) %>%
    # Keep only rows where the `names` column matches the genes of interest
    filter(names %in% genes) %>%
    
    # Create the plot
    ggplot(aes(x = names, y = logfoldchanges, fill = group)) +
      geom_bar(stat = "identity", position = position_dodge()) +
      xlab("genes") +
      ylab("LFC vs linker only") +
      scale_fill_brewer(palette = "Dark2")
}
```

## Number of Interactions

```{r, fig.height=5, fig.width=10}
interactionsLfc %>%
  filter(adj_pval < 0.1) %>%
  filter(abs(lfc) > 0.25) %>%
  count(interaction, name = "count") %>%  # Count occurrences for each interaction
  arrange(desc(count)) %>%  # Arrange data by decreasing counts
  mutate(interaction = factor(interaction, levels = interaction)) %>%
  ggplot(aes(x = interaction, y = count)) +
    geom_bar(stat = "identity") +
    ylab("# interaction DEGs") +
    ggtitle("Interaction DEG # (padj < 0.1, lfc > 0.25)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, fig.height=5, fig.width=5}
# Calculate total DEGs for individual ligands
totalDegs <- singlesLfc %>%
  filter(adj_pval < 0.1) %>%
  filter(abs(lfc) > 0.25) %>%
  select(name, interaction) %>%
  unique() %>%
  group_by(interaction) %>%
  summarise(total_degs = n())

# Count significant interactions
interactionCounts <- interactionsLfc %>%
  filter(adj_pval < 0.1) %>%
  filter(abs(lfc) > 0.25) %>%
  count(interaction, name = "count")

# Combine the two datasets
combinedData <- interactionCounts %>%
  full_join(totalDegs, by = "interaction") %>%
  mutate(interaction = factor(interaction, levels = interaction))  # Maintain original order

# Create the XY plot
ggplot(combinedData, aes(x = total_degs, y = count, label = interaction)) +
  geom_point(size = 2, alpha = 0.7) +  # Plot points
  geom_text_repel(size = 2) +
  xlim(0,4100) +
  labs(
    x = "Total Unique DEGs (Single Ligands)", 
    y = "Interaction DEGs", 
    title = "Total DEGs vs Interaction DEGs"
  )
```

Somewhat unsurprisingly the number of interaction DEGs scales with the total
number of unique DEGs in the corresponding single ligand stimulations. Notably
though there **is variance** in the number of interaction DEGs (e.g. the TNF
interactions on the right hand side) which is somewhat reassuring that the
interaction effects isn't just an artifact of noise in the data.

## Distribution of interaction effects

One interesting question is whether the same genes have interaction effects
across different ligands or whether, for the same ligand, the same genes tend to
promiscuously have interaction effects when paired with many different ligands.
We'll examine these one at a time

### Interaction genes across ligands

```{r}
# get number of times gene is interacting across all ligands
interactGeneSummary <- interactionsLfc %>%
  filter(adj_pval < 0.01) %>%
  filter(abs(lfc) > 0.25) %>%
  group_by(name) %>%
  summarise(num_interactions = n(),
            median_lfc = median(abs(lfc)),
            median_pval = median(adj_pval))
```

```{r, fig.height=5}
ggplot(interactGeneSummary, aes(x = num_interactions)) +
  geom_histogram(bins = 20) +
  ggtitle("Distribution of Interaction Effects Across Ligands")
```

```{r, fig.height=10, fig.width=14}
library(ggpointdensity)
p1 <- ggplot(interactGeneSummary, aes(x = factor(num_interactions), y=abs(median_lfc))) +
  geom_point() + 
  scale_color_viridis_c() +
  ylab("Median LFC (abs value)") +
  xlab("number of interactions")
p2 <- ggplot(interactGeneSummary, aes(x = factor(num_interactions), y=-log10(median_pval))) +
  geom_point() + 
  scale_color_viridis_c() +
  ylab("Median -log10(padj)") +
  xlab("number of interactions")
p3 <- ggplot(interactGeneSummary, aes(x = factor(num_interactions), y=abs(median_lfc))) +
  geom_boxplot() +
  ylab("Median LFC (abs value)") +
  xlab("number of interactions")
p4 <- ggplot(interactGeneSummary, aes(x = factor(num_interactions), y=-log10(median_pval))) +
  geom_boxplot() +
  ylab("Median -log10(padj)") +
  xlab("number of interactions")

p1+p2+p3+p4+plot_layout(ncol=2)
```

-   Majority of genes with interaction effects are present in a relatively small
    number of interactions (perhaps just that they are somewhat ligand-specific
    and each ligand has 7 potential pairings)
-   Genes with the highest median LFC and significance tend to be those with
    lower numbers of interactions, although genes with high numbers of
    interactions tend to have higher overall LFC and significance.
-   A small subset of genes with promiscuous interaction effects are highly
    significant (see below).

```{r}
# get promiscuous interactors with high significance
interactGeneSummary %>%
  filter(num_interactions > 10 & -log10(median_pval) > 20)
```

```{r}
# get promiscuous interactors with high significance
interactGeneSummary %>%
  filter(median_lfc > 1) %>%
  arrange(median_pval, ascending=T)
```

**Random interesting relatively unique interactions just from scanning**

-   IL6 + TNF -\> emergent Rorc upregulation (consistent with studies suggesting
    the role of Tnf in promoting Th17)
    -   <https://www.pnas.org/doi/10.1073/pnas.2109972118>
-   IL4 & IL2 + TNF -\> synergistic downregulation of Irf4 (role in Th1
    induction?)
-   IL4 + TNF -\> emergent Thrb upregulation
-   IFNA & IL27 + TNF -\> Cxcl10 upregulation
    -   IL27 also leads to synergistic Tbx21 induction but not IFNA

```{r, fig.height=10, fig.width=10}
p1 <- plot_lfc_pair(c("Rorc"),"IL6_TNF", degWilcox)
p2 <- plot_lfc_pair(c("Thrb","Snn","Irf4","Mdk"),"IL4_TNF", degWilcox)
p3 <- plot_lfc_pair(c("Irf4"),"IL2_TNF", degWilcox)
p6 <- plot_lfc_pair(c("Cxcl10", "Tbx21"),"IFNA_TNF", degWilcox)
p7 <- plot_lfc_pair(c("Cxcl10","Tbx21"),"IL27_TNF", degWilcox)

(p1+p3)/(p6+p7)/p2
```

### Interacting genes for each ligand

For the DEGs present induced by each ligand alone, let's see how many of them
have interaction effects and with how many of the pairs do they have effects.
The maximum number of pairs is 7 for any given ligand.

One thing to keep in mind with this analysis is I'm using DEGs derived from
wilcoxon test for the single stimulations and DEGs from the GLM model for the
interaction effects. Perhaps there are some differences in pvalue distributions
between these?

```{r}
pair_single_double_deg <- function(deg, interactions, lfc_thresh = 0.25, pval_thresh = 0.01) {
  # LFC and pval filter both the single ligand DEGs and the interaction DEGs
  
  # Create tibble with only singles
  degSingles <- deg %>%
    filter(grepl("_linker", group)) %>%
    mutate(group = gsub("_linker", "", group)) %>%
    filter(pvals_adj < pval_thresh & abs(logfoldchanges) > lfc_thresh)
  
  # Create tibble with number of significant interaction effects per gene per ligand
  groupInteractSummary <- interactions %>%
    separate(interaction, into = c("interaction1", "interaction2"), sep = "_") %>%
    pivot_longer(cols = c(interaction1, interaction2), 
                 names_to = "interaction_type", 
                 values_to = "group") %>%
    select(-interaction_type) %>%
    filter(adj_pval < pval_thresh & abs(lfc) > lfc_thresh) %>%
    group_by(group, name) %>%
    summarise(num_interactions = n(),
              median_interaction_lfc = median(lfc),
              median_interaction_padj = median(adj_pval), .groups = "drop")
  
  # Merge dataset on interactions to single DEGs
  degSinglesInteract <- degSingles %>% 
    left_join(groupInteractSummary, by = c("group" = "group", "names" = "name")) %>%
    mutate(num_interactions = replace_na(num_interactions, 0))
  
  # Calculate summary statistics per group per number of interactions
  degSummarySI <- degSinglesInteract %>%
    group_by(group, num_interactions) %>%
    summarise(interact_count = n(), .groups = "drop") %>%
    group_by(group) %>%
    mutate(interact_freq = interact_count / sum(interact_count)) %>%
    mutate(num_interactions = as_factor(num_interactions))
  
  out <- vector(mode = "list", length = 2)
  out[[1]] <- degSinglesInteract
  out[[2]] <- degSummarySI
  return(out)
}
```

```{r, fig.height=6, fig.width=15}
singleDoubleSig <- pair_single_double_deg(deg = degWilcox,
                                             interactions = interactionsLfc,
                                             lfc_thresh = 0,
                                             pval_thresh = 0.1)
singleDoubleSig[[2]]

# make plots
p1 <- ggplot(singleDoubleSig[[2]], aes(x=group,y=interact_count,fill=num_interactions)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_viridis_d(option = "turbo") +
  xlab("ligand") +
  ylab("# of DEGs\n in single ligand") +
  guides(fill=guide_legend(title="# of pairs\nwith interactions")) +
  ggtitle("Interacting genes per ligand\n(padj < 0.1)")
p2 <- ggplot(singleDoubleSig[[2]], aes(x=group,y=interact_freq,fill=num_interactions)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_viridis_d(option = "turbo") +
  xlab("ligand") +
  ylab("frequency of DEGs\n in single ligand") +
  guides(fill=guide_legend(title="# of pairs\nwith interactions")) +
  ggtitle("Interacting genes per ligand\n(padj < 0.1)")

p1+p2
```

```{r, fig.height=6, fig.width=15}
singleDoubleSigStrict <- pair_single_double_deg(deg = degWilcox,
                                             interactions = interactionsLfc,
                                             lfc_thresh = 0.25,
                                             pval_thresh = 0.01)
singleDoubleSigStrict[[2]]

# make plots
p1 <- ggplot(singleDoubleSigStrict[[2]], aes(x=group,y=interact_count,fill=num_interactions)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_viridis_d(option = "turbo") +
  xlab("ligand") +
  ylab("# of DEGs\n in single ligand") +
  guides(fill=guide_legend(title="# of pairs\nwith interactions")) +
  ggtitle("Interacting genes per ligand\n(lfc > 0.25, padj < 0.01)")
p2 <- ggplot(singleDoubleSigStrict[[2]], aes(x=group,y=interact_freq,fill=num_interactions)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_viridis_d(option = "turbo") +
  xlab("ligand") +
  ylab("frequency of DEGs\n in single ligand") +
  guides(fill=guide_legend(title="# of pairs\nwith interactions")) +
  ggtitle("Interacting genes per ligand\n(lfc > 0.25, padj < 0.01)")

p1+p2
```

```{r, fig.height=6, fig.width=15}
singleDoubleSigStrict2 <- pair_single_double_deg(deg = degWilcox,
                                             interactions = interactionsLfc,
                                             lfc_thresh = 1,
                                             pval_thresh = 0.01)
singleDoubleSigStrict2[[2]]

# make plots
p1 <- ggplot(singleDoubleSigStrict2[[2]], aes(x=group,y=interact_count,fill=num_interactions)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_viridis_d(option = "turbo") +
  xlab("ligand") +
  ylab("# of DEGs\n in single ligand") +
  guides(fill=guide_legend(title="# of pairs\nwith interactions")) +
  ggtitle("Interacting genes per ligand\n(lfc > 1, padj < 0.01)")
p2 <- ggplot(singleDoubleSigStrict2[[2]], aes(x=group,y=interact_freq,fill=num_interactions)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_viridis_d(option = "turbo") +
  xlab("ligand") +
  ylab("frequency of DEGs\n in single ligand") +
  guides(fill=guide_legend(title="# of pairs\nwith interactions")) +
  ggtitle("Interacting genes per ligand\n(lfc > 1, padj < 0.01)")

p1+p2
```

```{r, fig.height=8, fig.width=15}
singleDoubleSig[[1]] %>%
  ggplot(aes(x = as_factor(num_interactions),
             y = logfoldchanges, fill = as_factor(num_interactions))) +
    geom_boxplot() +
    scale_fill_viridis_d(option = "viridis") +
    facet_wrap(~group, ncol = 4) +
    ylab("LFC in single ligand condition") +
    xlab("# of pairs with interactions") +
    ggtitle("Relationship of # of interactions with LFC\n(padj < 0.1)") +
    theme(legend.position = "none")
```

```{r, fig.height=8, fig.width=15}
singleDoubleSigStrict[[1]] %>%
  ggplot(aes(x = as_factor(num_interactions),
             y = logfoldchanges, fill = as_factor(num_interactions))) +
    geom_boxplot() +
    scale_fill_viridis_d(option = "viridis") +
    facet_wrap(~group, ncol = 4) +
    ylab("LFC in single ligand condition") +
    xlab("# of pairs with interactions") +
    ggtitle("Relationship of # of interactions with LFC\n(lfc > 1, padj < 0.01)") +
    theme(legend.position = "none")
```

-   A good proportion of DEGs from a given ligand have some interaction,
    although this varies based on the exact cutoff you use (stricter lfc and
    padj cutoff = fewer percentage of DEGs have interactions)
    -   Given the relative uniqueness of IFNA genes, it's somewhat expected to
        see that there are in general fewer genes with interactions for this
        ligand
    -   IL-12 is interesting because at strict cutoffs, all of its DEGs have
        interaction effects with at least one other ligand. This is notable
        since IL-12 is a weird one (doesn't induce much DEGs)
-   A general trend is that the genes with higher numbers of interacting pairs
    for a given ligand have higher LFC in the single ligand condition. I guess
    this is reassuring but it also perhaps speaks to the fact that most
    interactions are these sort of antagonistic interactions between strong DEGs
    shared by two ligands.
    -   Interestingly this trend is stronger for certain ligands than for
        others. For example, IFNA has many genes with strong DEG and no
        interactions but for IL27 and IL21, the correlation between \# of pairs
        with interactions and LFC for a given gene is very strong

## Directionality of interaction effects

```{r, fig.height=8, fig.width=18}
singleDoubleSig[[1]] %>%
  mutate(median_interaction_lfc = replace_na(median_interaction_lfc,0)) %>%
  ggplot(aes(x = logfoldchanges, y = median_interaction_lfc)) +
    geom_pointdensity() +
    scale_color_viridis_c() +
    facet_wrap(~group, ncol = 4) +
    geom_hline(yintercept=0, linetype="dashed") +
    geom_vline(xintercept=0, linetype="dashed") +
    ylab("median LFC in interaction") +
    xlab("LFC in single ligand condition") +
    ggtitle("Relationship of interaction and single LFC\n(padj < 0.1)")
```

```{r, fig.height=8, fig.width=18}
singleDoubleSigStrict[[1]] %>%
  mutate(median_interaction_lfc = replace_na(median_interaction_lfc,0)) %>%
  ggplot(aes(x = logfoldchanges, y = median_interaction_lfc)) +
    geom_pointdensity() +
    scale_color_viridis_c() +
    facet_wrap(~group, ncol = 4) +
    geom_hline(yintercept=0, linetype="dashed") +
    geom_vline(xintercept=0, linetype="dashed") +
    ylab("median LFC in interaction") +
    xlab("LFC in single ligand condition") +
    ggtitle("Relationship of interaction and single LFC\n(padj < 0.1 & lfc > 0.25)")
```

It's clear that there's some directional bias. Of note, some interaction ligands
have abs(LFC) \< cutoff because the median function takes the mean when there's
only two samples. To get a better sense of this we can focus on the interaction
effects mapped onto the single-ligand effects for each ligand pair.

```{r generate_pairwise_plot_grid}
library(patchwork)

# Uses make_xy_plot function to generate a grid of plots with all pairwise interactions
generate_pairwise_plot_grid <- function(unique_ligands, interactions, singles, p_val_cutoff = 0.1, lfc_cutoff = 0.25) {
  
  # Create an empty list to store the plots
  plot_list <- list()
  
  # create tibble with significant iteraction DEGs
  interSig <- interactions %>%
    filter(adj_pval < p_val_cutoff)
  
  # Iterate through ligand pairs and generate plots
  for (row_ligand in unique_ligands) {
    for (col_ligand in unique_ligands) {
      # For diagonal plots, make them empty
      if (row_ligand == col_ligand) {
        plot_list[[paste(row_ligand, col_ligand, sep = "_")]] <- ggplot() + 
          theme_void()
        next
      }
      
      # Determine the ligand pair
      ligand_pair <- paste(row_ligand, col_ligand, sep = "_")
      reverse_pair <- paste(col_ligand, row_ligand, sep = "_")
      
      if (ligand_pair %in% interSig$interaction) {
        # Plot for the top-right side
        plot <- make_xy_plot(
          ligand_pair = ligand_pair,
          interactions = interactions,
          singles = singles,
          p_val_cutoff = p_val_cutoff,
          lfc_cutoff = lfc_cutoff
        )
        # Remove text and labels
        plot <- plot +
          theme(legend.position = "none")
        plot_list[[paste(row_ligand, col_ligand, sep = "_")]] <- plot
      } else if (reverse_pair %in% interSig$interaction) {
        # Transpose the plot for the bottom-left side
        plot <- make_xy_plot(
          ligand_pair = reverse_pair,
          interactions = interactions,
          singles = singles,
          p_val_cutoff = p_val_cutoff,
          lfc_cutoff = lfc_cutoff
        )
        # Remove text and labels, and transpose axes
        plot <- plot +
          coord_flip() + # Transpose the plot
          theme(legend.position = "none")
        plot_list[[paste(row_ligand, col_ligand, sep = "_")]] <- plot
      } else {
        # Add an empty plot if no interaction exists
        plot_list[[paste(row_ligand, col_ligand, sep = "_")]] <- ggplot() + 
          theme_void()
      }
    }
  }
  
  # Arrange the plots in a symmetric grid
  plot_grid <- wrap_plots(plot_list,
                          nrow = length(unique_ligands),
                          ncol = length(unique_ligands)) +
    plot_layout(heights = rep(3, length(unique_ligands)),
                widths = rep(3, length(unique_ligands)))
  
  # Return the plot grid
  return(plot_grid)
}
```

```{r, fig.height=30, fig.width=30}
# Extract unique ligands
unique_ligands <- unique(unlist(strsplit(interSig$interaction, "_")))

# Call the function
pairwisePlot <- generate_pairwise_plot_grid(
  unique_ligands = unique_ligands,
  interactions = interactionsLfc,
  singles = singlesLfc,
  p_val_cutoff = 0.1,
  lfc_cutoff = 0.25
)
pairwisePlot
#ggsave("plots/interactions_xy_plot_glmGamPoi.pdf")
```

## Strong emergent effects

We can try to identify "strong" interaction effects by focusing on those with
the largest difference between the interaction effect coefficient and the single
ligand coefficients from the GLM model.

```{r}
# get mean coeff for single ligands per interaction per gene
meanSingleCoeff <- singlesLfc %>%
  group_by(interaction, name) %>%
  summarise(mean_lfc_single = mean(lfc),
            max_abs_lfc_single = max(abs(lfc)),
            .groups = "drop")

interSigMerge <- interSig %>%
  select(interaction, name, adj_pval, lfc) %>%
  left_join(meanSingleCoeff, by = c("interaction","name")) %>%
  mutate(coeff_diff = lfc + mean_lfc_single) %>%
  arrange(coeff_diff)

interSigMerge
write_csv(emergentInter,"analysis_outs/interactions_diff_annotated.csv")
```

### Identifying "true" emergent interactions

To identify true "emergent" interaction effects, I will filter on the maximum
absolute value LFC in the singles.

```{r}
emergentInter <- interSigMerge %>%
  filter(max_abs_lfc_single < 0.1 & abs(lfc) > 0.4) %>%
  arrange(adj_pval)
emergentInter
write_csv(emergentInter,"analysis_outs/emergent_interactions_manual_cutoff.csv")
```

```{r}
plot_lfc_pair(c("Ramp1"),"IL2_TNF", degWilcox)
```

### Specific interesting cases

#### IL12 interaction effects

```{r, fig.height=18, fig.width=12}
p1 <- make_xy_plot("IL12_IFNA",
             interactions = interactionsLfc, singles = singlesLfc, p_val_cutoff=0.1) +
  geom_text_repel(aes(label=name), size=2) +
  facet_wrap(~lfc < 0, labeller = as_labeller(c(`TRUE` = "neg interaction LFC", `FALSE` = "pos interaction LFC")))
p2 <- make_xy_plot("IL12_IL27",
             interactions = interactionsLfc, singles = singlesLfc, p_val_cutoff=0.1) +
  geom_text_repel(aes(label=name), size=2) +
  facet_wrap(~lfc < 0, labeller = as_labeller(c(`TRUE` = "neg interaction LFC", `FALSE` = "pos interaction LFC")))
p3 <- make_xy_plot("IL12_IL21",
             interactions = interactionsLfc, singles = singlesLfc, p_val_cutoff=0.1) +
  geom_text_repel(aes(label=name), size=2) +
  facet_wrap(~lfc < 0, labeller = as_labeller(c(`TRUE` = "neg interaction LFC", `FALSE` = "pos interaction LFC")))
p1+p2+p3+plot_layout(ncol=1)
```

#### IL4 IFNA interaction effect

```{r, fig.height=6, fig.width=12}
make_xy_plot("IL4_IFNA",
             interactions = interactionsLfc, singles = singlesLfc, p_val_cutoff=0.1) +
  geom_text_repel(aes(label=name), size=2) +
  facet_wrap(~lfc < 0, labeller = as_labeller(c(`TRUE` = "neg interaction LFC", `FALSE` = "pos interaction LFC")))
```

#### IL4 vs IL6/IL21 interaction effects

```{r, fig.height=12, fig.width=12}
p1 <- make_xy_plot("IL4_IL6",
             interactions = interactionsLfc, singles = singlesLfc, p_val_cutoff=0.1) +
  geom_text_repel(aes(label=name), size=2) +
  facet_wrap(~lfc < 0, labeller = as_labeller(c(`TRUE` = "neg interaction LFC", `FALSE` = "pos interaction LFC")))
p2 <- make_xy_plot("IL4_IL21",
             interactions = interactionsLfc, singles = singlesLfc, p_val_cutoff=0.1) +
  geom_text_repel(aes(label=name), size=2) +
  facet_wrap(~lfc < 0, labeller = as_labeller(c(`TRUE` = "neg interaction LFC", `FALSE` = "pos interaction LFC")))
p1/p2
```
